---
title: "Non Hierarchical Tennis Model 1 v2"
format:
  html:
    toc: true
    code-tools: true
    code-line-numbers: true
    number-sections: true
    mainfont: Georgia, serif
editor: source
---

## 

```{r}
if (!require(tidybayes)) {
    install.packages("tidybayes")
    library(tidybayes)
}

if (!require(tidyr)) {
    install.packages("tidyr")
    library(tidyr)
}

if (!require(brms)) {
    install.packages("brms")
    library(brms)
}

if(!require(ggplot2)) {
  install.packages("ggplot2")
  library(ggplot2)
}

if(!require(reliabilitydiag)) {
  install.packages("reliabilitydiag")
  library(reliabilitydiag)
}


if (!require(metadat)) {
  install.packages("metadat")
  library(metadat)
}

if(!require(cmdstanr)){
    install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
    library(cmdstanr)
}
cmdstan_installed <- function(){
  res <- try(out <- cmdstanr::cmdstan_path(), silent = TRUE)
  !inherits(res, "try-error")
}
if(!cmdstan_installed()){
    install_cmdstan()
}
```

```{r}
data <- read.csv("../clean-data/master-sub.csv")

head(data)
```

Extract relevant data

```{r}
df <- data[c("ServerWon", "BreakPoint", "Tiebreaker")]

str(df$ServerWon)
str(df$BreakPoint)
str(df$Tiebreaker)

# mean(na.omit(df$ServerWon))
```

Fit model

```{r}

fit <- brm(
    formula = ServerWon ~ BreakPoint + Tiebreaker,
    data = df,
    family = bernoulli(link = "logit"),
    prior = c(
      prior(normal(0,2^2), class = "Intercept"),
      prior(normal(0,1^2), class = "b")
    ),
    chains = 4,
    iter = 2000,
    cores = 4
)

```

Checking results

```{r}

summary(fit)

```

Win probabilities

```{r}

post <- spread_draws(fit, b_Intercept, b_BreakPoint, b_Tiebreaker)

p_alpha <- inv_logit_scaled(post$b_Intercept)
p_bp <- inv_logit_scaled(post$b_Intercept+post$b_BreakPoint)
p_tb <- inv_logit_scaled(post$b_Intercept+post$b_Tiebreaker)

mean(p_alpha)
mean(p_bp)
mean(p_tb)



library(tidybayes)
library(dplyr)
library(ggplot2)

post <- spread_draws(fit, b_Intercept, b_BreakPoint, b_Tiebreaker) %>%
  mutate(
    p_regular  = plogis(b_Intercept),
    p_break    = plogis(b_Intercept + b_BreakPoint),
    p_tiebreak = plogis(b_Intercept + b_Tiebreaker)
  ) %>%
  select(.draw, p_regular, p_break, p_tiebreak) %>%
  pivot_longer(-.draw, names_to = "situation", values_to = "p")

ggplot(post, aes(x = p, y = situation)) +
  stat_pointinterval(.width = c(.5, .8, .95)) +
  xlim(0, 1) +
  labs(x = "P(server wins point)", y = NULL)
```

## MCMC Plots

```{r}
library(brms)

# trace plots
mcmc_plot(fit,
          pars = c("b_Intercept", "b_BreakPoint", "b_Tiebreaker"),
          type = "trace")

# overlaid posterior densities by chain
mcmc_plot(fit,
          pars = c("b_Intercept", "b_BreakPoint", "b_Tiebreaker"),
          type = "dens_overlay")

```

## Posterior predictive plots

```{r}
newdata <- data.frame(
  Tiebreaker = 0,
  BreakPoint = 1
)

ppd_break <- posterior_epred(
  fit,
  newdata = newdata
  )

ppd_reg <- posterior_epred(
  fit,
  newdata = df[sample(nrow(df), 1),]
)

ppd_tie <- posterior_epred(
  fit,
  newdata = data.frame(
    Tiebreaker = 1,
    BreakPoint = 0
  )
)

preds <- data.frame(
  p_reg = as.vector(ppd_reg),
  p_break = as.vector(ppd_break),
  p_tie = as.vector(ppd_tie)
)
```

```{r}
p <- ggplot(preds) + 
  geom_density(
    aes(p_reg, colour = 'Any point')
  ) + geom_density(
    aes(p_break, colour = 'Breakpoint')
  ) + geom_density(
    aes(p_tie, colour = 'Tiebreak')
  ) +
  xlab('Win probability') +
  ylab('Density')
p
```

## Posterior predictive plots 2 (using posterior_predict)

```{r}
# Posterior predictive draws of the *outcome* (0/1), not just p
yrep <- posterior_predict(fit, newdata = df, ndraws = 500)

```

```{r}
y_obs_bar  <- mean(df$ServerWon)
yrep_bar   <- rowMeans(yrep)

hist(yrep_bar, breaks = 40,
     main = "Posterior predictive distribution of mean serve win rate",
     xlab = "Mean ServerWon in replicated data")
abline(v = y_obs_bar, col = "red", lwd = 2)
```

## Posterior predictive checks by situation

```{r}
is_reg  <- df$BreakPoint == 0 & df$Tiebreaker == 0
is_bp   <- df$BreakPoint == 1
is_tb   <- df$Tiebreaker == 1

# Observed proportions
obs_reg  <- mean(df$ServerWon[is_reg])
obs_bp   <- mean(df$ServerWon[is_bp])
obs_tb   <- mean(df$ServerWon[is_tb])

# replicated proportions for each posterior draw
rep_reg <- rowMeans(yrep[, is_reg, drop = FALSE])
rep_bp  <- rowMeans(yrep[, is_bp,  drop = FALSE])
rep_tb  <- rowMeans(yrep[, is_tb,  drop = FALSE])

# simple base R plots
par(mfrow = c(1, 3))

hist(rep_reg, breaks = 30,
     col = "lightskyblue2",
     main = "Regular points",
     xlab = "Mean ServerWon (rep)")
abline(v = obs_reg, col = "red", lwd = 2)

hist(rep_bp, breaks = 30,
     col = "lightskyblue2",
     main = "Break points",
     xlab = "Mean ServerWon (rep)")
abline(v = obs_bp, col = "red", lwd = 2)

hist(rep_tb, breaks = 30,
     col = "lightskyblue2",
     main = "Tiebreak points",
     xlab = "Mean ServerWon (rep)")
abline(v = obs_tb, col = "red", lwd = 2)

par(mfrow = c(1, 1))  # reset
```

## Additional convergence diagnostics

Sampler diagnostics

```{r}
np <- nuts_params(fit)
head(np)
```

Divergences

```{r}
divergences <- subset(np, Parameter == "divergent__" & Value == 1)
n_div <- nrow(divergences)
n_div
```

Max treedepth hits:

```{r}
treedepth_hits <- subset(np, Parameter == "treedepth__" & Value == max(Value))
nrow(treedepth_hits)
```

## Reliabilitydiag

```{r}
library(brms)
library(reliabilitydiag)

# posterior mean coefficients
b <- fixef(fit)  # rows: "Intercept", "BreakPoint", "Tiebreaker"

b0   <- b["Intercept",  "Estimate"]
b_bp <- b["BreakPoint", "Estimate"]
b_tb <- b["Tiebreaker","Estimate"]

# predicted win prob for each point (plug-in logistic regression)
linpred   <- b0 + b_bp * df$BreakPoint + b_tb * df$Tiebreaker
fit_probs <- plogis(linpred)   # length = nrow(df)

# reliability diagram for all points
rd_all <- reliabilitydiag(fit_probs, y = df$ServerWon)
autoplot(rd_all, type = "miscalibration") +
  coord_cartesian(xlim = c(0.58, 0.7))
```

## ROC Plot

```{r}
## ROC Plot

# install/load pROC if needed
if (!require(pROC)) {
  install.packages("pROC")
  library(pROC)
}

# posterior mean (plug-in) predictions: P(ServerWon = 1 | BreakPoint, Tiebreaker)
b <- fixef(fit)  # rows: "Intercept", "BreakPoint", "Tiebreaker"

b0   <- b["Intercept",  "Estimate"]
b_bp <- b["BreakPoint", "Estimate"]
b_tb <- b["Tiebreaker", "Estimate"]

linpred   <- b0 + b_bp * df$BreakPoint + b_tb * df$Tiebreaker
fit_probs <- plogis(linpred)  # predicted probabilities for each point

# ROC object
roc_obj <- roc(response = df$ServerWon,
               predictor = fit_probs,
               quiet = TRUE)

# base ROC plot with AUC
plot(roc_obj,
     main = "ROC curve for non-hierarchical model",
     xlab = "False positive rate (1 - specificity)",
     ylab = "True positive rate (sensitivity)")
text(0.6, 0.1,
     labels = paste0("AUC = ", round(auc(roc_obj), 3)))

```

## PSIS–LOO

```{r}
loo_fit <- loo(fit, ndraws = 1000)
loo_fit
```

## LOO-PIT-ECDF

```{r}
pp_check(
fit,
type = "loo_pit_ecdf", # LOO-PIT ECDF calibration check
ndraws = 500 # you can adjust if you like
)
```

## Sensitivity Analysis (3 other scenarios)

```{r}
library(brms)

## three alternative priors

prior_narrow <- c(
  prior(normal(0, 1^2),   class = "Intercept"),
  prior(normal(0, 0.5^2), class = "b")
)

prior_wide <- c(
  prior(normal(0, 5^2), class = "Intercept"),
  prior(normal(0, 2^2), class = "b")
)


# Center the intercept at p0 ≈ 0.615
p0     <- 0.615
mu_int <- qlogis(p0)   # logit(0.615)
#mu_int =~ 0.468

prior_inform <- c(
  prior(normal(0.468, 0.2^2), class = "Intercept"),
  prior(normal(0,      0.4^2), class = "b")
)

## refit models with the new priors

fit_narrow <- brm(
  formula = ServerWon ~ BreakPoint + Tiebreaker,
  data    = df,
  family  = bernoulli(link = "logit"),
  prior   = prior_narrow,
  chains  = 4,
  iter    = 2000,
  cores   = 4
)

fit_wide <- brm(
  formula = ServerWon ~ BreakPoint + Tiebreaker,
  data    = df,
  family  = bernoulli(link = "logit"),
  prior   = prior_wide,
  chains  = 4,
  iter    = 2000,
  cores   = 4
)


fit_inform <- brm(
  formula = ServerWon ~ BreakPoint + Tiebreaker,
  data    = df,
  family  = bernoulli(link = "logit"),
  prior   = prior_inform,
  chains  = 4,
  iter    = 2000,
  cores   = 4
)

```

#### Summaries of the four fits

```{r}
summary(fit)
summary(fit_narrow)
summary(fit_wide)
summary(fit_inform)
```

#### Plot win probabilities

```{r}
library(tidybayes)
library(dplyr)
library(tidyr)
library(ggplot2)

post_base <- spread_draws(fit, b_Intercept, b_BreakPoint, b_Tiebreaker) %>%
  mutate(model = "original")

post_narrow <- spread_draws(fit_narrow, b_Intercept, b_BreakPoint, b_Tiebreaker) %>%
  mutate(model = "narrow")

post_wide <- spread_draws(fit_wide, b_Intercept, b_BreakPoint, b_Tiebreaker) %>%
  mutate(model = "wide")

post_inform <- spread_draws(fit_inform, b_Intercept, b_BreakPoint, b_Tiebreaker) %>%
  mutate(model = "informative")

post_all <- bind_rows(post_base, post_narrow, post_wide, post_inform) %>%
  mutate(
    p_regular  = plogis(b_Intercept),
    p_break    = plogis(b_Intercept + b_BreakPoint),
    p_tiebreak = plogis(b_Intercept + b_Tiebreaker)
  ) %>%
  select(.draw, model, p_regular, p_break, p_tiebreak) %>%
  pivot_longer(cols = starts_with("p_"),
               names_to = "situation",
               values_to = "p")

# numeric comparison
post_all %>%
  group_by(model, situation) %>%
  summarise(
    mean = mean(p),
    l95  = quantile(p, 0.025),
    u95  = quantile(p, 0.975),
    .groups = "drop"
  )

# simple plot
ggplot(post_all, aes(x = p, y = situation, colour = model)) +
  stat_pointinterval(.width = c(0.5, 0.95),
                     position = position_dodge(width = 0.4)) +
  xlim(0, 1) +
  labs(x = "P(server wins point)", y = NULL,
       title = "Sensitivity to prior choice (including informative prior)")

```

#### Plot posterior densities for each

```{r}
library(dplyr)
library(ggplot2)

# ensure a consistent order of scenarios in the legend
post_all$model <- factor(post_all$model,
                         levels = c("original", "narrow", "wide", "informative"))

# Any point
post_reg <- filter(post_all, situation == "p_regular")
ggplot(post_reg, aes(x = p, colour = model)) +
  geom_density() +
  labs(x = "Win probability",
       y = "Density",
       colour = "Scenario",
       title = "Any point")

# Break point
post_break <- filter(post_all, situation == "p_break")
ggplot(post_break, aes(x = p, colour = model)) +
  geom_density() +
  labs(x = "Win probability",
       y = "Density",
       colour = "Scenario",
       title = "Break point")

# Tiebreak
post_tie <- filter(post_all, situation == "p_tiebreak")
ggplot(post_tie, aes(x = p, colour = model)) +
  geom_density() +
  labs(x = "Win probability",
       y = "Density",
       colour = "Scenario",
       title = "Tiebreak")

```

```{r}
library(dplyr)
library(ggplot2)

post_all$model <- factor(post_all$model,
                         levels = c("original", "narrow", "wide", "informative"))

post_all_plot <- post_all %>%
  filter(situation %in% c("p_regular", "p_break", "p_tiebreak")) %>%
  mutate(
    situation = recode(
      situation,
      p_regular = "Any point",
      p_break   = "Breakpoint",
      p_tiebreak = "Tiebreak"
    )
  )

ggplot(post_all_plot, aes(x = p, colour = model)) +
  geom_density() +
  facet_wrap(~ situation, nrow = 1, scales = "free") +
  labs(
    x = "Win probability",
    y = "Density",
    colour = "Scenario"
  )

```
